---
title: "Class 10"
output: html_notebook
---

```{r}
library(tidyverse)
```

# Review regression, minimizing mean squared error, finding line of best fit

```{r}
d = read_csv("../pset2/english.csv")
d$RT = d$RTlexdec
```

# What is R-squared? Proportion of variance explained

```{r}
```

# Introducing the regression lm()

```{r}
```


# What is the Std.Error? Measures uncertainty 

```{r}
l = lm(data=d, RT ~ Familiarity)
d$Resid = residuals(l)
sqrt(sum(d$Resid^2)/((nrow(d) - 2) * sum((d$Familiarity - mean(d$Familiarity))^2)))

a = NULL
for (i in 1:1000) {
  d.new = sample_n(d, nrow(d), replace=T)
  l = lm(data=d.new, RT ~ Familiarity)
  fam.coef = coef(l)[2]
  a = append(a, fam.coef)
}

hist(a, breaks=50)
sd(a)
summary(l)

-.063/sd(a)


```

# Now scramble the values and see what the coefficient is

```{r}
```

# Now using a continuous variable as a predictor (WrittenFrequency), see how well you can predict RT

```{r}

```

# Which is better at predicting RT? Familiarity or WrittenFrequency?

```{r}

```


