---
title: "Class 8 - Uniforms and Normals"
output: html_notebook
---

```{r}
library(tidyverse)
x = c(1, 5, 10, 3, NA)
mean(x, na.rm=T)
sd(x)
sd(x, na.rm=T)
```

# Review from Class 7

Review mean, median, standard deviation, normal distributions.

```{r}
rnorm(5, 0, 1)

mean(rnorm(5, 0, 1))

mean(rnorm(100000, 0, 1))

sd(rnorm(100000, 0, 1))

sd(rnorm(5, 0, 1))


rbinom(3, 1, .5)

x = rnorm(100000, 0, 1)
mean(x < 0)
mean(-1 < x & x < 1)

#
runif(5, 5, 10)
runif(5, -12, 20)
hist(runif(1000, -12, 20))
```


# Exercise 2


```{r}
d = read_csv("../textbook/data/perry_winter_2017_iconicity.csv")
arrange(d, -Iconicity) %>%
  select(Word, POS, Iconicity)

arrange(d, Iconicity) %>%
  select(Word, POS, Iconicity)
```


Get the iconicity scores from d above. Rank them, look at them.

Now make a histogram. Find the mean and standard deviation.

```{r, width=3, height=3}
hist(d$Freq, breaks=50)
hist(d$Iconicity)

icon.mean = mean(d$Iconicity)
icon.sd = sd(d$Iconicity)
icon.nrow = nrow(d)
```

Now sample a normal distribution with that mean and standard deviation. 
How similar are they?

```{r}
print(hist(rnorm(icon.nrow, icon.mean, icon.sd), breaks=50))
print(hist(d$Iconicity, breaks=50))

icon = tibble(icon=d$Iconicity, type="real")
sampled = tibble(icon=rnorm(icon.nrow, icon.mean, icon.sd), type="sim")
icon.sampled = bind_rows(icon, sampled)
ggplot(icon.sampled, aes(x=icon, fill=type)) +
  geom_histogram(alpha=.4)  +
  facet_grid(type ~ .)

nrow(filter(icon.sampled, type == "real"))
nrow(filter(icon.sampled, type == "sim"))

```

# Mean as a model

Task: for a random word, predict the iconicity.

Idea: Use the mean.

```{r}
d$Prediction = mean(d$Iconicity)
```

How wrong will we be!? (I.e.: what is the error?)

```{r}
d$Residual = d$Iconicity - d$Prediction
sum(d$Residual) # this is close to 0 because we have a mixture of negatives and positives
sum(abs(d$Residual)) # absolute total error
sum(abs(d$Residual))/nrow(d)
```

How wrong will we be (squared)?

Instead of absolute difference, look at squared difference.

Mean Squared Error is a crucial quantity for linear regression and statistics.

```{r}
d$SquareDiff = d$Residual * d$Residual
hist(d$SquareDiff)
sum(d$SquareDiff)
sum(d$SquareDiff)/nrow(d)
```

Now do it for POS I.e., compare POS == "Interjection" vs. POS == "Name"

Use Mean Squared Error to assess how wrong we are.

Using group means minimizes least square error MORE than the overall mean.

```{r}
group_by(d, POS) %>%
  summarise(icon.mean=mean(Iconicity))

d = group_by(d,POS) %>%
  mutate(pos.mean = mean(Iconicity))

d$Residual.Pos = d$pos.mean - d$Iconicity
d$SquareDiff.Pos = d$Residual.Pos * d$Residual.Pos

# using overall mean
sum(d$SquareDiff)/nrow(d)

# using group means
sum(d$SquareDiff.Pos)/nrow(d)
```

# Overfitting with coins

```{r}
rbinom(3, 1, .5)
```
