---
title: "Class 7 - Uniforms and Normals"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
d = read_csv("../textbook/data/perry_winter_2017_iconicity.csv")
arrange(d, -Iconicity)
```

# Understanding mean, median, and standard deviation (sd)

```{r}
x = c(1, 2, 3, 4, 5)
y = c(1, 2, 3, 4, 100000000)

mean(x)
mean(y)

median(x)
median(y)
```
# Standard deviation

On average, how far is a data point from the mean?

```{r}
print(x)
mean(abs(x - 3))
sum(abs(x - 3))/4
sqrt(var(x))

sqrt(sum((x - 3)**2)/4)

x
y
sd(x)
sd(y)
```

# Uniform and normal

```{r, fig.height=3, fig.width=3}
u = runif(10000, 0, 1)
hist(u)

u1 = runif(3, 0, 1)
u2 = runif(3, 0, 1)
u3 = runif(3, 0, 1)
print(u1)
print(u2)

n = rnorm(10000, 0, 1)
hist(n)

d = tibble(unif=u, norm=n)
```

# Make some different histograms by taking the mean of the uniform distribution. 

```{r, fig.height=3, fig.width=3}
get_unif_mean = function(n) {
  return(mean(runif(n, 0, 1)))
}

a = NULL
for (i in 1:1000) {
  a = append(a, get_unif_mean(10000))
}
hist(a)
```
```{r, fig.height=3, fig.width=3}
a = NULL
for (i in 1:1000) {
  a = append(a, get_unif_mean(5))
}
hist(a)


mean(runif(100000, 0, 1))
```

# Exercise

Take 10000 draws from a uniform distribution [0, 1]. Find the 2.5 percentile, 50th percentile, 97.5th percentile empirically (by computing it). That is: find the value for which 2.5% of the data is less than that value. Then for 50% (this is the median!), etc. 

```{r}
samples = tibble(u=runif(10000, 0, 1))
samples = arrange(samples, u)
samples[250, ]
samples[5000, ]
samples[9750, ]
hist(samples$u)
quantile(samples$u, probs = c(.025, .5, .975))
```

Let's try with a small sample size. 
```{r}
samples = tibble(u=runif(10, 0, 1))
samples = arrange(samples, u)
samples[250, ]
samples[5000, ]
samples[9750, ]
hist(samples$u)
quantile(samples$u, probs = c(.025, .5, .975))
```
Now do the same for a normal distribution. Keep the mean 0, but play around with the standard deviation (try 1, 2, 3). How does this affect things?

```{r}
x = rnorm(100000,0,1)
quantile(x,c(.025,.5,.975))

hist(x)
```

# Exercise 2

Get the iconicity scores from d above. Rank them, look at them.

Now make a histogram. Find the mean and standard deviation. 

Now sample a normal distribution with that mean and standard deviation. 
How similar are they?

```{r}
```