---
title: "Class 6-Dative and ggplot"
output: html_document
---
# What is a function?

```{r}
# input: filename (as a character string)
# output: dataframe in R
dative = read_csv("../pset1/dative.csv")
mean(dative$LengthOfRecipient)

add.s.to.verb = function(x) {
  return (paste(x, "s", sep=""))
}

add.s.to.verb("give")
dative$Verb.3S = add.s.to.verb(dative$Verb)

# LengthOfTheme as a function of LengthOfRecipient
ggplot(dative, aes(x=LengthOfRecipient, y=LengthOfTheme)) + 
  geom_point()


```


```{r}
ggplot(dative, aes(x=LengthOfRecipient, y=LengthOfTheme)) + 
  geom_point() + 
  geom_line()
```

### Nonpronominal theme
Amy gave the book to Greg. (PP)
Amy gave Greg the book. (NP)

### Pronominal theme
Amy gave it to Greg. (PP) S-theme-recipient
* Amy gave Greg it. (NP) S-recipient-theme
* Amy gave Greg who was coming back from the movies and had just 
seen Mank it.
Greg was given it by Amy. reicipient-theme-S

```{r}
head(dative)
unique(dative$PronomOfTheme)
group_by(dative, PronomOfTheme) %>%
  summarise(mean.NP = mean(RealizationOfRecipient == "NP"))

group_by(dative, PronomOfTheme) %>%
  summarise(mean.NP = mean(RealizationOfRecipient == "NP")) %>%
  ggplot(data=., aes(x=PronomOfTheme, y=mean.NP)) + 
  geom_bar(stat="identity")
```

Confound: something which offers an alternative explanation for the observed pattern in the data, which cannot be ruled out.

Based on the above, it's tempting to say: "When you have a Pronoun theme, you avoid NP." BUT....we also know that when you have a short theme and a long recipient, you are more likely to put the recipient last. 

* Amy gave Greg it. (NP) S-recipient-theme
* Amy gave [Greg who was coming back from the movies and had just 
seen Mank] it.

```{r fig.width=6, fig.height=3}
dative.pro.len = group_by(dative, PronomOfTheme, LengthOfTheme) %>%
  summarise(mean.NP = mean(RealizationOfRecipient == "NP"),
            n=n())

ggplot(dative.pro.len, aes(x=LengthOfTheme,
                           mean.NP,
                           color=PronomOfTheme)) + 
  geom_point() + 
  geom_line()
```


```{r}
ggplot(dative.pro.len, aes(x=LengthOfTheme,
                           y=mean.NP,
                           color=PronomOfTheme,
                           size=n)) + 
  geom_point() 
```

Cluster data points together, when they have too few data points.

```{r}

nrow(filter(dative, LengthOfTheme > 15))
dative.pro.len = mutate(dative.pro.len, 
                        LengthOfTheme.Trunc = ifelse(LengthOfTheme >= 5,
                                                     5,
                                                     LengthOfTheme))
ggplot(dative.pro.len, aes(x=LengthOfTheme.Trunc,
                           y=mean.NP,
                           color=PronomOfTheme,
                           size=n)) + 
  geom_point() 
```


```{r}
dative.pro.len = group_by(dative, PronomOfTheme, PronomOfRec, LengthOfTheme) %>%
  summarise(mean.NP = mean(RealizationOfRecipient == "NP"),
            n=n())

ggplot(dative.pro.len, aes(x=LengthOfTheme,
                           mean.NP,
                           color=PronomOfTheme)) + 
  geom_point() +
  #facet_grid(ROW ~ COL)
  facet_grid(. ~ PronomOfRec)
```

WHen you have a nonpronominal recipient and a pronominal theme, you are very unlikely to get an NP realization.

* Amy gave Greg it. (NP) S-recipient-theme

It seems better when the recipient is also a pronoun.

* Amy gave him it.

```{r}
ggplot(dative.pro.len, aes(x=LengthOfTheme,
                           mean.NP,
                           color=PronomOfTheme)) + 
  geom_point() +
  #facet_grid(ROW ~ COL)
  facet_grid(PronomOfRec ~ PronomOfTheme)
```