---
title: "Class 7 - Uniforms and Normals"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
d = read_csv("../textbook/data/perry_winter_2017_iconicity.csv")
arrange(d, -Iconicity)
```

# Understanding mean, median, and standard deviation (sd)

```{r}
x = c(1, 2, 3, 4, 5)
y = c(1, 2, 3, 4, 100000000)
```

# Uniform and normal

```{r}
u = runif(10000, 0, 1)
n = rnorm(10000, 0, 1)

d = tibble(unif=u, norm=n)
```

# Make some different histograms by taking the mean of the uniform distribution. 

```{r}
```

# Exercise

Take 10000 draws from a uniform distribution [0, 1]. Find the 2.5 percentile, 50th percentile, 97.5th percentile empirically (by computing it). That is: find the value for which 2.5% of the data is less than that value. Then for 50% (this is the median!), etc. 

Now do the same for a normal distribution. Keep the mean 0, but play around with the standard deviation (try 1, 2, 3). How does this affect things?

```{r}
```

# Exercise 2

Get the iconicity scores from d above. Rank them, look at them.

Now make a histogram. Find the mean and standard deviation. 

Now sample a normal distribution with that mean and standard deviation. 

How similar are they?

```{r}
```