---
title: "Class 14"
output: html_notebook
---

```{r}
library(tidyverse)
```

# Back to our favorite data set: Lexical decision times

```{r}
d = read_csv("../pset2/english.csv")
d$RT = d$RTlexdec
```

# Multiple regression

```{r}
ggplot(d, aes(x=LengthInLetters, y=RT)) + geom_point()

group_by(d, LengthInLetters) %>%
  summarise(m=mean(RT))

l = lm(RT ~ LengthInLetters, data=d)
summary(l)

l.log = lm(RT ~ log(LengthInLetters), data=d)
summary(l.log)


l2 = lm(RT ~ LengthInLetters + WrittenFrequency, data=d)

# y = mx + b -> one predictor
# y = beta1 * x1 + beta2 * x2 + b -> two predictors
```

Predict the RT of a word which has 5 letters and a WrittenFrequency of 6.91 (in log space).

```{r}
l.5 = lm(data=d, RT ~ WrittenFrequency + LengthInLetters +
           Ncount + MeanBigramFrequency + WrittenSpokenFrequencyRatio)
summary(l.5)
```

# Categorical predictors

```{r}
unique(d$AgeSubject)

d$Predict.Mean = mean(d$RT)
d = group_by(d, AgeSubject) %>%
  mutate(Predict.AgeGroup = mean(RT))
unique(d$Predict.AgeGroup)

sse.mean.as.model = sum((d$RT - d$Predict.Mean)^2)
sse.agegroup.as.model = sum((d$RT - d$Predict.AgeGroup)^2)

1 - sse.agegroup.as.model/sse.mean.as.model

d$AgeVariable = ifelse(d$AgeSubject == "young", 0, 1)
sample_n(d, 10) %>% select(AgeSubject, AgeVariable)

l.cat = lm(data=d, RT ~ AgeVariable)
# l.cat = lm(data=d, RT ~ 1 + AgeVariable)
# l.cat = lm(data=d, RT ~ 0 + AgeVariable) # get rid of intercept

e = tibble(AgeVariable = 0)

predict(l.cat, newdata=e)

summary(l.cat)
# y = mx + b
# young -> predict intercept -> 6.43
# old -> predict intercept + 1 * slope -> 6.43 + .22 -> 6.65

unique(d$Predict.AgeGroup)
```

# Combine multiple regression with categorical predictors

```{r}
l.cat.mult = lm(data=d, RT ~ WrittenFrequency + 
                  LengthInLetters + AgeVariable)
summary(l.cat.mult)

l.cat.mult.2 = lm(data=d, RT ~ WrittenFrequency + 
                  LengthInLetters + AgeSubject)
summary(l.cat.mult.2)
```

# More interpretable predictors by normalizing

```{r}
d$WrittenFrequency.Normalized = (d$WrittenFrequency - mean(d$WrittenFrequency))/sd(d$WrittenFrequency)

d$LengthInLetters.Normalized = (d$LengthInLetters - mean(d$LengthInLetters))/sd(d$LengthInLetters)

d$AgeVariable.Normalized = (d$AgeVariable - mean(d$AgeVariable))/sd(d$AgeVariable)

l.cat.mult.3 = lm(data=d, RT ~ WrittenFrequency.Normalized + 
                  LengthInLetters.Normalized + AgeVariable.Normalized)

```

# Testing the regression assumptions

## Are the residuals normal?

## Does the variance of the residuals change? Heteroscedasticity

## Adjusted R2

## Review: can you compute R2 manually for RT ~ WrittenFrequency + LengthInLetteres?
```{r}
l.simple = lm(data=d, RT ~ WrittenFrequency + LengthInLetters)
```
## Adjusted R2: 

```{r}

```